The effect of prenatal environment on brain metabolic function and perception in a lizard

### 2025-09-06 ms.qmd
updates

#### Statistical analyses
We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and detection latency) separately. We first fit a set of preliminary models for all the interest variables (detection latency, mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction. Models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables). For detection latency, we also included previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation as fixed effects. We included the interaction between CORT and motivation because CORT can impact appetite [@conde2018stress]. All models included clutch identity as a random factor to account for clutch effects. For models of detection latency we also included a random effect of lizard identity because of our repeated measures design. The structure and results of these models are provided on Tables S1-S12 of the Supplementary Material. After the preliminary models, we simplified models by re-fitting to exclude factors that were not significant. However, our models always included the main interest factors: CORT, temperature and their interaction. Random factors remain the same for the final models. All the response variables were mean centered and standardized by dividing by two times the standard deviation [@gelman2008scaling]. Before standardization, mitochondrial density, DNA damage, lipid peroxidation, and detection latency were log-transformed.

@@

To further explore the relationships between mitochondrial physiology and detection latency, we used a Structural Equation Modelling (SEM) approach. We fit a multivariate brm from stan [@stan] in R (version 4.4.0) [@R] for each brain region/stimulus separately. We included in the model all the variables of interest and their interactions structured following a specific set of hypotheses (see @fig-sem_results_OB and @fig-sem_results_OT). Because experience with prey did not affect lizards’ behaviour (see Tables S11, S12), to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors. Clutch identity was included as a random factor for all variables. For our SEM model, the error structure was modelled assuming a Gaussian distribution and we estimated residual correlations among all variables.

@@

```{r, models_mitochondrial}
#| label: models_mitochondrial
#
# Run models mitochondrial physiology (each region separately)
#
var_m <- c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros", "mean_dnadamage", "mean_peroxidation")
regions <- c("OB", "OT")
for (p in var_m){
  for (h in regions){
    if (h == "OB"){
      df <- clean_df %>% filter(region == "OB")
      l <- "OB"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    } else {
      df <- clean_df %>% filter(region == "OT")
      l <- "OT"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    }

@@

**____________________FROM HERE____________________**
```{r, organise_posteriors}
#| label: organise_posteriors
#
# Organising the posteriors of the previous models to fit the tables below
#
source(here("R", "func.R"))
#
# Databases for each region/stimulus
post_OB <- data.frame()
post_OT <- data.frame()
# Names posteriors:
region <- c("OB", "OT")
names_OB <- c("m_def_mean_mitodensity_OB", "m_def_mean_potential_OB", "m_def_mean_ros_OB", "m_def_mean_dnadamage_OB", "m_def_mean_peroxidation_OB", "m_def_t_D_Chemical")
names_OT <- c("m_def_mean_mitodensity_OT", "m_def_mean_potential_OT", "m_def_mean_ros_OT", "m_def_mean_dnadamage_OT", "m_def_mean_peroxidation_OT", "m_def_t_D_Visual")
#
# Organising the results
for (r in region) {
  model_select <- get(paste0("names_", r))   
  for (pos in model_select) {
    model <- get(pos)      # Get the model from the global environment
    post_result <- tidy_post(model)        # Apply tidy_post to each model
    # Add a new column to identify the region and model
    post_result$Region <- r
    post_result$Model <- pos
    # Append to the appropriate data frame
    if (r == "OB") {
      post_OB <- bind_rows(post_OB, post_result)
    } else {
      post_OT <- bind_rows(post_OT, post_result)
    }
  }
}
```

```{r, values_posteriors}
#| label: values_posteriors
#
# Extracting the posteriors for the models and the values of interest. Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
#
source(here("R", "func.R"))
#
# A) Olfactory bulbs/Chemical stimulus
#
MD_OB <- post_values(m_def_mean_mitodensity_OB, "none")
MP_OB <- post_values(m_def_mean_potential_OB, "none")
ROS_OB <- post_values(m_def_mean_ros_OB, "none")
DNA_OB <- post_values(m_def_mean_dnadamage_OB, "none")
LP_OB <- post_values(m_def_mean_peroxidation_OB, "none")
DET_OB <- post_values(m_def_t_D_Chemical, "none")
#
# B) Optic tecta/Visual stimulus
MD_OT <- post_values(m_def_mean_mitodensity_OT, "none")
MP_OT <- post_values(m_def_mean_potential_OT, "none")
ROS_OT <- post_values(m_def_mean_ros_OT, "none")
DNA_OT <- post_values(m_def_mean_dnadamage_OT, "sex")
LP_OT <- post_values(m_def_mean_peroxidation_OT, "none")
DET_OT <- post_values(m_def_t_D_Visual, "none")
```

#### Does prenatal CORT and temperature affect mitochondrial function and oxidative stress in the brain?
Models explained between 37.8 to 50.4 % of the variation in mitochondrial function and 39.9 to 59 % for oxidative stress (see Table S1).

_Olfactory bulbs_: age significantly increased DNA damage and lipid peroxidation (Table S3). However, we did not find any significant effects of CORT, temperature, or their interaction on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2).

```{r, fig-results_energy}
#| label: fig-results_energy
#| fig-cap: "Estimates of mitochondrial density (A, C) and metabolic capacity (B, D) in the olfactory bulbs (A, B) and optic tecta (C, D) of L. delicata hatchlings as a function of the different prenatal conditions. Black dots indicate the posterior mean, and the bars represent the SD of the estimates. The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions. Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
#| fig-name: "fig-results_energy"
#
source(here("R", "func.R"))
#
# A) Plotting the results for OB/Chemical stimulus using the df from before
plot_density_OB <- plotting(MD_OB, "Mit density")
plot_potential_OB <- plotting(MP_OB, "Mit potential")
#
fig_OB_energy <- plot_grid(plot_density_OB, NULL, plot_potential_OB, NULL,
                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
#
# B) Plotting the results for OT/Visual stimulus
plot_density_OT <- plotting(MD_OT, "Mit density")
plot_potential_OT <- plotting(MP_OT, "Mit potential")
#
fig_OT_energy <- plot_grid(plot_density_OT, NULL, plot_potential_OT, NULL,
                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
#
# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
plot_legend_bottom <- plotting(MP_OB, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
legend_plot <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
# C) Merging plots
#
# Create the figure grid with extra space for images
fig_results_energy <- plot_grid(
  fig_OB_energy, fig_OT_energy, NULL,
  nrow = 3, rel_heights = c(0.9, 0.9, 0.15))
# Final composition: Merge everything, adding images and legend
final_plot_energy <- ggdraw(fig_results_energy) +
  # Insert OB image in the top right
  draw_image("./Others/OB.png", x = 0.87, y = 0.79, width = 0.12, height = 0.2) +
  # Insert OT image in the middle right
  draw_image("./Others/OT.png", x = 0.87, y = 0.34, width = 0.11, height = 0.18) +
  # Insert legend at the bottom-right
  draw_grob(legend_plot, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001) +
  # Insert title for each plot
  annotate("text", x = 0.032, y = 0.97, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.487, y = 0.97, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.032, y = 0.515, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.49, y = 0.515, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold")
#
# Print final plot
ggsave(here("./output/figures/text/results_energy.png"), plot = final_plot_energy, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_energy.png")
```

_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2). However, the interaction of CORT and temperature have significant effects on DNA damage in the optic tecta (see @fig-results_oxidative). For DNA damage, CORT elevations had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`), but it decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E). Prenatal CORT had opposite effects on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), but differences between groups were not significant (see @fig-results_oxidative C). In addition, we found that females had higher DNA damage levels, and lipid peroxidation increased with age (see Table S4).

@@

@@

fig_OB_oxidative_top <- plot_grid(plot_ros_OB, NULL, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, NULL, plot_peroxidation_OB,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))

@@

@@

fig_OT_oxidative_top <- plot_grid(plot_ros_OT, NULL, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, NULL, plot_peroxidation_OT,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))

@@

  fig_OB_oxidative, NULL, fig_OT_oxidative, NULL,
  nrow = 4, rel_heights = c(1, 0.05, 1, 0.1))

@@

  draw_image("./Others/OB.png", x = 0.56, y = 0.755, width = 0.2, height = 0.3) +

@@

  draw_image("./Others/OT.png", x = 0.56, y = 0.28, width = 0.19, height = 0.28) +
  # Insert legend at the bottom-right
  draw_grob(legend_plot, x = 0.5, y = 0.025, width = 0.0001, height = 0.0001) +

@@

  annotate("text", x = 0.0365, y = 0.986, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0365, y = 0.755, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.567, y = 0.755, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0365, y = 0.5, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0365, y = 0.268, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.567, y = 0.268, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("segment", x = 0.46, xend= 0.46, y = 0.21, yend = 0.26, size = 0.5, colour = "black") +
  annotate("segment", x = 0.45, xend= 0.46, y = 0.21, yend = 0.21, size = 0.5, colour = "black") +
  annotate("segment", x = 0.45, xend= 0.46, y = 0.26, yend = 0.26, size = 0.5, colour = "black") +
  annotate("text", x = 0.47, y = 0.228, label = "*", hjust = 0.5, vjust = 0.5, size = 7)

@@

ggsave(here("./output/figures/text/results_oxidative.png"), plot = final_plot_oxidative, width = 18, height = 24, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_oxidative.png")

@@

#### Does prenatal CORT and incubation temperature affect prey detection?
Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).

@@

var <- c("MD", "MP", "ROS", "DNA", "LP", "DET")

@@

@@

@@

  mutate(Variable = factor(Variable, levels = c("Mit density", "Mit potential", "ROS", "DNA damage", "Lipid peroxidation", "Detection latency"))) %>%

@@

  flextable::compose(i = c(2:18,20:36), j = 1, value = as_paragraph(""), part = "body") %>%
  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24,26,27,29,30,32,33,35,36),

>>>>

We recorded five physiological variables from our flow cytometry assay: mitochondrial density, metabolic capacity, ROS production, DNA damage, and lipid peroxidation. These variables reflect average per-cell values for each sample, as fluorescence intensity was normalized by the number of gated events (cells) recorded. However, to assess how early environmental conditions may influence mitochondrial function at the organelle level, we created two additional response variables by dividing metabolic capacity and ROS production by mitochondrial density. This procedure assumes that mitochondrial density serves as a proxy for the number of mitochondria per cell, and the derived variables provide an estimate of the average activity per mitochondrion, therefore allowing us to assess mitochondrial-specific responses to early-life conditions.

#### Statistical analyses
We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, metabolic capatity/mitochondrial density, ROS, ROS/mitochondrial density, DNA damage, lipid peroxidation, and detection latency) separately. We first fit a set of preliminary models for all the interest variables where we included the hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction. These models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables). For detection latency, we also included previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation as fixed effects. We included the interaction between CORT and motivation because CORT can impact appetite [@conde2018stress]. All models included clutch identity as a random factor to account for clutch effects. For models of detection latency we also included a random effect of lizard identity because of our repeated measures design. The structure and results of these models are provided on Tables S1-S22 in _Supplementary Material_. After the preliminary models, we simplified models by excluding some of the factors that were not significant. However, we always included the main interest factors: CORT, temperature and their interaction. Random factors remain the same for the final models. All the response variables were mean centered and standardized by dividing by two times the standard deviation [@gelman2008scaling]. Before standardization, mitochondrial density, DNA damage, lipid peroxidation, and detection latency were log-transformed.

@@

To further explore the relationships between mitochondrial physiology and detection latency, we used a Structural Equation Modelling (SEM) approach. We fit a multivariate brm from stan [@stan] in R (version 4.4.0) [@R] for each brain region/stimulus separately. We included in the model all the variables of interest and their interactions structured following a specific set of hypotheses (see @fig-sem_results_OB and @fig-sem_results_OT). Because experience with prey did not affect lizards’ behaviour (see Tables S21, S22), to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors. Clutch identity was included as a random factor for all variables. For our SEM model, the error structure was modelled assuming a Gaussian distribution and we estimated residual correlations among all variables.

@@

```{r, models_mitochondrial}
#| label: models_mitochondrial
#
# Run models mitochondrial physiology (each region separately)
#
var_m <- c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros", "mean_dnadamage", "mean_peroxidation")
regions <- c("OB", "OT")
for (p in var_m){
  for (h in regions){
    if (h == "OB"){
      df <- clean_df %>% filter(region == "OB")
      l <- "OB"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    } else {
      df <- clean_df %>% filter(region == "OT")
      l <- "OT"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    }

@@

```{r, organise_posteriors}
#| label: organise_posteriors
#
# Organising the posteriors of the previous models to fit the tables below
#
source(here("R", "func.R"))
#
# Databases for each region/stimulus
post_OB <- data.frame()
post_OT <- data.frame()
# Names posteriors:
region <- c("OB", "OT")
names_OB <- c("m_def_mean_mitodensity_OB", "m_def_mean_potential_OB", "m_def_mean_conpotential_OB", "m_def_mean_ros_OB", "m_def_mean_conros_OB", "m_def_mean_dnadamage_OB", "m_def_mean_peroxidation_OB", "m_def_t_D_Chemical")
names_OT <- c("m_def_mean_mitodensity_OT", "m_def_mean_potential_OT", "m_def_mean_conpotential_OT", "m_def_mean_ros_OT", "m_def_mean_conros_OT", "m_def_mean_dnadamage_OT", "m_def_mean_peroxidation_OT", "m_def_t_D_Visual")
#
# Organising the results
for (r in region) {
  model_select <- get(paste0("names_", r))   
  for (pos in model_select) {
    model <- get(pos)      # Get the model from the global environment
    post_result <- tidy_post(model)        # Apply tidy_post to each model
    # Add a new column to identify the region and model
    post_result$Region <- r
    post_result$Model <- pos
    # Append to the appropriate data frame
    if (r == "OB") {
      post_OB <- bind_rows(post_OB, post_result)
    } else {
      post_OT <- bind_rows(post_OT, post_result)
    }
  }
}
```

```{r, values_posteriors}
#| label: values_posteriors
#
# Extracting the posteriors for the models and the values of interest. Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
#
source(here("R", "func.R"))
#
# A) Olfactory bulbs/Chemical stimulus
#
MD_OB <- post_values(m_def_mean_mitodensity_OB, "none")
MP_OB <- post_values(m_def_mean_potential_OB, "none")
MPcon_OB <- post_values(m_def_mean_conpotential_OB, "none")
ROS_OB <- post_values(m_def_mean_ros_OB, "none")
ROScon_OB <- post_values(m_def_mean_conros_OB, "none")
DNA_OB <- post_values(m_def_mean_dnadamage_OB, "none")
LP_OB <- post_values(m_def_mean_peroxidation_OB, "none")
DET_OB <- post_values(m_def_t_D_Chemical, "none")
#
# B) Optic tecta/Visual stimulus
MD_OT <- post_values(m_def_mean_mitodensity_OT, "none")
MP_OT <- post_values(m_def_mean_potential_OT, "none")
MPcon_OT <- post_values(m_def_mean_conpotential_OT, "none")
ROS_OT <- post_values(m_def_mean_ros_OT, "none")
ROScon_OT <- post_values(m_def_mean_conros_OT, "none")
DNA_OT <- post_values(m_def_mean_dnadamage_OT, "sex")
LP_OT <- post_values(m_def_mean_peroxidation_OT, "none")
DET_OT <- post_values(m_def_t_D_Visual, "none")
```

#### Do prenatal CORT and temperature affect mitochondrial function and oxidative stress in the brain?
Models explained between 37.8 to 50.4 % of the variation in mitochondrial function and 39.9 to 59 % for oxidative stress (see Table S1).

_Olfactory bulbs_: age significantly increased DNA damage and lipid peroxidation (Table S3). However, we did not find any significant effects of CORT, temperature on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2). The posterior distribution provided little support for an overall CORT-temperature interaction on lipid peroxidation ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OB$Control_Hot) - mean(LP_OB$CORT_Hot)) - (mean(LP_OB$Control_Cold) - mean(LP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OB$Control_Hot - LP_OB$CORT_Hot) - (LP_OB$Control_Cold - LP_OB$CORT_Cold)), 3, equal = TRUE)`). However, targeted contrasts showed that lipid peroxidation was lower in CORT-Hot compared to CORT-Cold lizards (β~CORT-Hot~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OB$CORT_Hot) - mean(LP_OB$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OB$CORT_Hot - LP_OB$CORT_Cold), 3, equal = TRUE)`).

```{r, fig-results_energy}
#| label: fig-results_energy
#| fig-cap: "Estimates of mitochondrial density (A, C) and metabolic capacity (B, D) in the olfactory bulbs (A, B) and optic tecta (C, D) of L. delicata hatchlings as a function of the different prenatal conditions. Black dots indicate the posterior mean, and the bars represent the SD of the estimates. The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions. Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
#| fig-name: "fig-results_energy"
#
source(here("R", "func.R"))
#
# A) Plotting the results for OB/Chemical stimulus using the df from before
plot_density_OB <- plotting(MD_OB, "Mit density")
plot_potential_OB <- plotting(MP_OB, "Mit potential")
plot_conpotential_OB <- plotting(MPcon_OB, "Mit potential/Mit density")
#
fig_OB_energy <- plot_grid(plot_density_OB, NULL, 
                          plot_potential_OB, plot_conpotential_OB,
                          nrow = 2, ncol = 2, rel_widths = c(1))
#
# B) Plotting the results for OT/Visual stimulus
plot_density_OT <- plotting(MD_OT, "Mit density")
plot_potential_OT <- plotting(MP_OT, "Mit potential")
plot_conpotential_OT <- plotting(MPcon_OT, "Mit potential/Mit density")
#
fig_OT_energy <- plot_grid(plot_density_OB, NULL, 
                          plot_potential_OB, plot_conpotential_OB,
                          nrow = 2, ncol = 2, rel_widths = c(1))
#
# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
plot_legend_bottom <- plotting(MP_OB, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
legend_plot <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
# C) Merging plots
#
# Create the figure grid with extra space for images
fig_results_energy <- plot_grid(
  fig_OB_energy, fig_OT_energy, legend_plot,
  nrow = 3, rel_heights = c(1, 1, 0.1))
# Final composition: Merge everything, adding images and legend
final_plot_energy <- ggdraw(fig_results_energy) +
  # Insert OB image in the top right
  draw_image("./Others/OB.png", x = 0.75, y = 0.75, width = 0.21, height = 0.3) +
  # Insert OT image in the middle right
  draw_image("./Others/OT.png", x = 0.75, y = 0.28, width = 0.19, height = 0.3) +
  # Insert title for each plot
  annotate("text", x = 0.035, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.035, y = 0.755, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.535, y = 0.755, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.035, y = 0.52, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.035, y = 0.29, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.535, y = 0.29, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold")

#
# Print final plot
ggsave(here("./output/figures/text/results_energy.tiff"), plot = final_plot_energy, width = 18, height = 24, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_energy.tiff")
```

_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2). However, the interaction of CORT and temperature have significant effects on DNA damage ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot)) - (mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OT$Control_Hot - DNA_OT$CORT_Hot) - (DNA_OT$Control_Cold - DNA_OT$CORT_Cold)), 3, equal = TRUE)`), and lipid peroxidation ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot)) - (mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OT$Control_Hot - LP_OT$CORT_Hot) - (LP_OT$Control_Cold - LP_OT$CORT_Cold)), 3, equal = TRUE)`) in the optic tecta. CORT elevations had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`), but it decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E). The direction of the effects of CORT on lipid peroxidation were the opposite on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), with significant differences between CORT-Hot and CORT-Cold lizards (β~CORT-Hot~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$CORT_Hot) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$CORT_Hot - LP_OT$CORT_Cold), 3, equal = TRUE)`; see @fig-results_oxidative C). In addition, we found that females had higher DNA damage levels, and lipid peroxidation increased with age (see Table S4).

@@

plot_conros_OB <- plotting(ROScon_OB, "ROS/Mit density")

@@

fig_OB_oxidative_top <- plot_grid(plot_ros_OB, plot_conros_OB, NULL,
                    nrow = 1, rel_widths = c(1, 1, 0.35))
fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, plot_peroxidation_OB, NULL,
                    nrow = 1, rel_widths = c(1, 1, 0.35))

@@

plot_conros_OT <- plotting(ROScon_OT, "ROS/Mit density")

@@

fig_OT_oxidative_top <- plot_grid(plot_ros_OT, plot_conros_OT, NULL,
                    nrow = 1, rel_widths = c(1, 1, 0.35))
fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, plot_peroxidation_OT, NULL,
                    nrow = 1, rel_widths = c(1, 1, 0.35))

@@

  fig_OB_oxidative, fig_OT_oxidative, legend_plot,
  nrow = 3, rel_heights = c(1, 1, 0.1))

@@

  draw_image("./Others/OB.png", x = 0.85, y = 0.77, width = 0.17, height = 0.24) +

@@

  draw_image("./Others/OT.png", x = 0.85, y = 0.295, width = 0.15, height = 0.245) +

@@

  annotate("text", x = 0.028, y = 0.986, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.455, y = 0.986, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.028, y = 0.745, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.455, y = 0.745, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.028, y = 0.51, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.455, y = 0.51, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.028, y = 0.27, label = "G", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.455, y = 0.27, label = "H", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("segment", x = 0.39, xend= 0.39, y = 0.217, yend = 0.262, size = 0.5, colour = "black") +
  annotate("segment", x = 0.38, xend= 0.39, y = 0.217, yend = 0.217, size = 0.5, colour = "black") +
  annotate("segment", x = 0.38, xend= 0.39, y = 0.262, yend = 0.262, size = 0.5, colour = "black") +
  annotate("text", x = 0.4, y = 0.24, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
  annotate("segment", x = 0.83, xend= 0.83, y = 0.173, yend = 0.262, size = 0.5, colour = "black") +
  annotate("segment", x = 0.82, xend= 0.83, y = 0.173, yend = 0.173, size = 0.5, colour = "black") +
  annotate("segment", x = 0.82, xend= 0.83, y = 0.262, yend = 0.262, size = 0.5, colour = "black") +
  annotate("text", x = 0.84, y = 0.21, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
  annotate("segment", x = 0.83, xend= 0.83, y = 0.648, yend = 0.737, size = 0.5, colour = "black") +
  annotate("segment", x = 0.82, xend= 0.83, y = 0.648, yend = 0.648, size = 0.5, colour = "black") +
  annotate("segment", x = 0.82, xend= 0.83, y = 0.737, yend = 0.737, size = 0.5, colour = "black") +
  annotate("text", x = 0.84, y = 0.685, label = "*", hjust = 0.5, vjust = 0.5, size = 7)

@@

ggsave(here("./output/figures/text/results_oxidative.tiff"), plot = final_plot_oxidative, width = 24, height = 24, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_oxidative.tiff")

@@

#### Do prenatal CORT and incubation temperature affect prey detection?
Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).

@@

var <- c("MD", "MP", "MPcon", "ROS", "ROScon", "DNA", "LP", "DET")

@@

      Variable == "MPcon" ~ "Mit potential/Mit density",

@@

      Variable == "ROScon" ~ "ROS/Mit density",

@@

  mutate(Variable = factor(Variable, levels = c("Mit density", "Mit potential", "Mit potential/Mit density", "ROS", "ROS/Mit density","DNA damage", "Lipid peroxidation", "Detection latency"))) %>%

@@

  flextable::compose(i = c(2:24,26:48), j = 1, value = as_paragraph(""), part = "body") %>%
  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24,26,27,29,30,32,33,35,36,38,39,41,42,44,45,47,48),
